# Reproducible Research: Peer Assessment 1
```{r chunkDefaults}
opts_chunk$set(echo=TRUE)
```
## Loading and preprocessing the data
```{r loadData}
# working directory is set to repo as RStudio project.
data <- read.csv(file="activity.csv") # load data
data$date <- as.Date(data$date) # convert format into date
```
## What is mean total number of steps taken per day?
```{r histogram}
# aggregate/sum the steps for each day
dataDay <- aggregate(steps~date,data=data, FUN = "sum")
# histogram for the total steps per day
hist(dataDay$steps, main="Histogram of total steps per day",xlab="steps per day")
# calculate mean
meanSteps <- mean(dataDay$steps)
# calculate median
medianSteps <- median(dataDay$steps)
```
The *mean* of the total steps taken per day is `r format(meanSteps, digit=2)`.  
The *median* of the total steps taken per day is `r medianSteps`.

## What is the average daily activity pattern?

```{r timeseries}
# calculate the average for each time intervall
dataInterval <- aggregate(steps~interval,data=data, FUN = "mean")
# make time sereis plot
plot(dataInterval$interval, dataInterval$steps, type="l", main="Time Series Plot", xlab="Interval", ylab="Average steps per interval")
# calculate the inverall with maximum averagesteps
maxStepsInt <- dataInterval[dataInterval$steps==max(dataInterval$steps),"interval"]
```

The interval `r maxStepsInt` has the maximum of average steps.

## Imputing missing values
```{r missingValues}
# calculate number of missing values
missValues <- table(is.na(data$steps))[2]
# manipulate data and replace NAs with mean
dataMan <- data
# number intervals per day
intDay <- nrow(subset(data,day="2012-10-01"))
dataMan[is.na(dataMan$steps),"steps"] <- meanSteps/intDay
# aggregate/sum the steps for each day
dataManDay <- aggregate(steps~date,data=dataMan, FUN = "sum")
# histogram for the total steps per day
hist(dataManDay$steps, main="Histogram of total steps per day",xlab="steps per day")
# calculate mean
meanStepsMan <- mean(dataManDay$steps)
# calculate median
medianStepsMan <- median(dataManDay$steps)
# difference in mean
meanStepsMan - meanSteps
# difference in median
medianStepsMan - medianSteps
```
The mean of the steps from the manipulated data set is `r format(meanStepsMan, digit=2)`.  
The median of the steps from the manipulated data set is `r format(medianStepsMan, digit=2)`.
## Are there differences in activity patterns between weekdays and weekends?
```{r weekdays}
# determine weekday
dataMan$day <- weekdays(dataMan$date)
# weekday data
dataIntervalWeekday <- aggregate(steps~interval,data=subset(dataMan, day %in% c("Monday","Tuesday", "Wednesday", "Thursday", "Friday")), FUN = "mean")
# weekend data
dataIntervalWeekend <- aggregate(steps~interval,data=subset(dataMan, day %in% c("Saturday","Sunday")), FUN = "mean")
# time series plots
plot(dataIntervalWeekday$interval, dataIntervalWeekday$steps, type="l", main="Time Series Plot - Weekday", xlab="Interval", ylab="Average steps per interval")
plot(dataIntervalWeekend$interval, dataIntervalWeekend$steps, type="l", main="Time Series Plot - Weekend", xlab="Interval", ylab="Average steps per interval")
```